---
title: "p8105_hw2_al3998"
author: "AimingLiu"
date: "9/24/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(dplyr)
library(ggplot2)
```

## Problem 1
```{r  message=FALSE}
#Import the data#
mrtrash_data = read_excel("~/Desktop/data science/p8105_hw2_al3998/data/TrashWheel.xlsx",
              sheet = "Mr. Trash Wheel",skip = 1)%>%
             janitor::clean_names()%>% 
             drop_na(dumpster)%>%
             select(dumpster:homes_powered) %>% 
             mutate(sports_balls = as.integer(round(sports_balls, digits = 0)))

preci_17_data = read_excel("~/Desktop/data science/p8105_hw2_al3998/data/TrashWheel.xlsx",
               sheet = "2017 Precipitation",range = "A2:B14")%>%
                janitor::clean_names()%>%
                drop_na(total)%>%
                mutate(year = "2017")             

preci_18_data = read_excel("~/Desktop/data science/p8105_hw2_al3998/data/TrashWheel.xlsx",
                sheet = "2018 Precipitation",range = "A2:B14")%>%
                janitor::clean_names()%>%
                drop_na(total)%>%
                mutate(year = "2018")
preci_data = bind_rows(preci_17_data,preci_18_data)%>%
   mutate(month = month.name[month])
```
After reading and cleaning the data in the Mr. Trash Wheel sheet,there are `r nrow(mrtrash_data)` obs.of `r ncol(mrtrash_data)` variables,the variable columns in the data set include `r mrtrash_data %>% select(dumpster:volume_cubic_yards) %>% colnames`and some other columns,among which all the numbers in the column "sports_balls" have been transformed into integer.

After reading and cleaning the data in the 2018 Precipitation,there are `r nrow(preci_18_data)`  obs.of `r ncol(preci_18_data)` variables including columns `r colnames(preci_18_data)`.
After reading and cleaning the data in the 2017 Precipitation,there are `r nrow(preci_17_data)` obs.of `r ncol(preci_17_data)`variables including columns of `r colnames(preci_17_data)`.

After  combining 2018 Precipitation and 2017 Precipitation, the data in the data set called preci_data,there are `r nrow(preci_data)` obs. of `r ncol(preci_data)` and include `r colnames(preci_data)`.

The total precipitation in 2018 is `r sum(preci_18_data %>% drop_na(month) %>%pull(total))`.
And the median of sports balls in a dumpster in 2017 is `r median((mrtrash_data)%>%filter(year=="2017")%>% pull(sports_balls))`.


## Problem 2
```{r message=FALSE}
pols_month_data = read_csv("~/Desktop/data science/p8105_hw2_al3998/data/pols-month.csv")%>%
                  janitor::clean_names()%>%
                  separate(mon,into = c("year", "month","day"), sep = "-")%>%
                  mutate(month = month.name[as.numeric(month)])%>%
                  pivot_longer(cols = starts_with("prez"),
                               names_to = "president",
                               names_prefix = "prez_" )%>%
                  filter(value != 0)%>%
                  select(-day,-value)
 
snp_data  =   read_csv("~/Desktop/data science/p8105_hw2_al3998/data/snp.csv")%>%               
              janitor::clean_names()%>%
              separate(date,into = c("month", "day","year"), sep = "/")%>%
              arrange(year, month)%>%
              mutate(month = month.name[as.numeric(month)])%>%
              select(year,month,close,-day)

 
 unemployment_data = read_csv("~/Desktop/data science/p8105_hw2_al3998/data/unemployment.csv")%>%
                     pivot_longer(
                     cols=Jan:Dec,
                     names_to = "month")%>%
                     arrange(Year,month)%>%
                     mutate(month = match(month, month.abb))%>%
                     mutate(month = month.name[month])%>%
                     mutate(year = as.character(Year))
## merging data                    
pols_snp_data = left_join(pols_month_data, snp_data,by = c("year","month"))      
pols_snp_une_data = left_join(pols_snp_data, unemployment_data,by = c( "year","month"))
```

The first data frame "pols_month_data" is a data frame with `r nrow(pols_month_data)` rows and `r ncol(pols_month_data)` columns, the range of the year in this adta set is from `r range(pols_month_data %>% pull(year))[1]` to `r range(pols_month_data %>% pull(year))[2]`,and the name of key variables in this data set are `r names(pols_month_data)`

The second data frame "snp_data" is a data frame with `r nrow(snp_data)` rows and `r ncol(snp_data)` columns, the range of the year in this adta set is from `r range(snp_data %>% pull(year))[1]` to `r range(snp_data %>% pull(year))[2]`,and the name of key variables in this data set are `r names(snp_data)`

The third data frame "unemployment_data" is a data frame with `r nrow(unemployment_data)` rows and `r ncol(unemployment_data)` columns, the range of the year in this adta set is from `r range(unemployment_data %>% pull(year))[1]` to `r range(unemployment_data %>% pull(year))[2]`,and the name of key variables in this data set are `r names(unemployment_data)`

After merging the three data sets,the data sets called "pols_snp_une_data" include `r nrow(pols_snp_une_data)` rows and `r ncol(pols_snp_une_data)` columns.Some data in this data set is missing because that the range of years are different.

## Problem 3
## Load and tidy the data
```{r  message=FALSE}
popular_baby_data = read_csv("~/Desktop/data science/p8105_hw2_al3998/data/Popular_Baby_Names.csv")%>%
                    janitor::clean_names()%>%
                    mutate(ethnicity = recode(ethnicity,"ASIAN AND PACIFIC ISLANDER" = "ASIAN AND PACI","BLACK NON HISPANIC" = "BLACK NON HISP","WHITE NON HISPANIC" = "WHITE NON HISP"))%>%
                     mutate(childs_first_name = str_to_title(childs_first_name))%>%
                     distinct()
```
## Produce a table showing the rank in popularity of the name “Olivia”
```{r}
    Olivia_data = popular_baby_data %>% 
    filter(gender == "FEMALE" & childs_first_name == "Olivia") %>%
    pivot_wider(names_from = year_of_birth, values_from = rank,id_cols = ethnicity)
```

## Producing a table showing the most popular name among male children over time.
```{r}
   male_data = filter(popular_baby_data,gender == "MALE",rank =="1")%>%
   select(ethnicity,year_of_birth,childs_first_name) %>% 
   pivot_wider(names_from = year_of_birth, values_from = childs_first_name)
```

## producing a scatter plot
```{r}
  plot_data = filter(popular_baby_data,gender == "MALE",ethnicity == "WHITE NON HISP",year_of_birth == "2016")
  ggplot(plot_data,aes(x = rank,y = count))+geom_point()+xlab( "rank of the names")+ ylab( "number of children with a name")
```
 
#When the rank of the names gets higher,there are more children have this name

